preprocessing

### 2022-11-21 r-prep/prep.Rmd
Integrated new energy data for deving countries

```{r}
load("data-raw/data_iea.RData")

data_kaya = data_base %>% 
  group_by(across(-c(sector_title,subsector_title))) %>% 
  summarise(CO2=sum(CO2,na.rm=TRUE)) %>%
  left_join(data_iea,by=c("ISO","year")) %>%
  mutate(energy_intensity = energy/GDP_pc) %>%
  mutate(carbon_intensity = CO2/energy) %>%
  na.omit() %>% 
  pivot_longer(c(CO2,pop,GDP_pc,energy_intensity,carbon_intensity),names_to = "var",values_to = "value")

data_kaya$value[0]<-NA

data_kaya%<>%
  na.omit()

>>>>

#load("data-raw/data_iea.RData") Thanks to Will we got better Data now. 
#load better energy data 
data_energy<-read_excel("data-raw/data_energy_co2_countries.xlsx",sheet = "data")
# filter for ISO, Year and Energy
data_energy%<>%
  select(ISO,year,energy)

data_kaya = data_base %>% 
  group_by(across(-c(sector_title,subsector_title,N2O, CH4,GHG,co2_lulucf,gdp_ppp, Fgas,GHG_pc,GHG_pg,CO2))) %>% 
  na.omit() %>% 
  summarise(CO2=sum(CO2,na.rm=T))%>%
  left_join(data_energy,by=c("ISO","year")) %>%
  mutate(energy_intensity = energy/GDP_pc) %>%
  mutate(carbon_intensity = CO2/energy) %>%
  na.omit() %>% 
  pivot_longer(c(CO2,pop,GDP_pc,energy_intensity,carbon_intensity),names_to = "var",values_to = "value")
